### 2020-01-23 Chapters/DataGeneration.tex

\chapter{Data generation and storage}

Explain about mono files plus ambisonics IRs.

\section{Recorded IRs}

AmbisonicsDRIR: explain problematics, standard, etc.
also: pysofaconventions

\section{Simulated IRs}

explain different methods and libraries.
tell about masp.

\section{High-level scene description}

ambiscaper