### 2020-01-23 Chapters/ScientificBackground.tex

\chapter{Scientific Background}

\section{State of the Art}