### 2020-03-04 Chapters/CoherenceEstimation.tex
update

\chapter{Coherence Estimation}

\section{Introduction}

A number of practical applications benefit of the knowledge about the diffuseness of a sound field, including speech enhancement and dereverberation \cite{p_habets_dual-microphone_2006}, noise suppression \cite{ito_designing_2010}, source separation \cite{duong_under-determined_2009} or background estimation \cite{stefanakis_foreground_2015}. In the field of spatial audio, diffuseness estimation is often used for parametrization \cite{pulkki_directional_2006, politis_compass_2018}, Direction-of-Arrival estimation \cite{thiergart_localization_2009} or source separation \cite{motlicek_real-time_2013}.

\todo{In this paper}, we study diffuseness estimation by subjecting a tetrahedral microphone array to spherically isotropic noise fields.
The motivation for this work is, first, that tetrahedral arrays are a well known type of microphone arrays, which have today become popular for applications related to Virtual and Augmented Reality. 
Second, the spherical isotropic sound field is known to be a good approximation to the reverberant part of the sound field in a room \cite{elko_spatial_2001, mccowan_microphone_2003}, and therefore it would be interesting to investigate how different microphone arrays behave under such conditions.

\subsection{\label{subsec:3:1} Coherence analysis}

Diffuseness is commonly estimated through the
\textit{Magnitude Squared Coherence} (MSC) \cite{elko_spatial_2001} between two frequency-domain signals $S_1$ and $S_2$, as a function of the
\textit{wavenumber} $k$ and the microphone distance $r$:
\begin{equation}
    \text{MSC}_{12}(k r) =
	\frac{|\left\langle S_1(k r) S_2(k r)^* \right\rangle|^2}
	{\left\langle|S_1(k r)|^2\right\rangle \left\langle|S_2(k r)|^2\right\rangle},
    \label{eq:MSC}
\end{equation}
where the $\left\langle \cdot \right\rangle$ operator represents the temporal
expected value, and $^*$ defines the complex conjugate operator. In the case of spherical isotropic noise fields, Eq.~(\ref{eq:MSC})
can be expressed in terms of microphone directivity patterns
$T(\phi,\theta,k r)$ as \cite{elko_spatial_2001}:

\begin{equation}
	\begin{aligned}
&\text{MSC}_{12}(k r) = \frac{|N_{12}(k r)|^2}{|D_{12}(kr)|^2} \\
&= \frac{|\int_{0}^{\pi} \int_{0}^{2\pi} T_1(\phi,\theta,k r) T_2^*(\phi,\theta,k r) e^{-jk r cos\theta} sin\theta d\theta d\phi|^2}{|\sqrt{ \int_{0}^{\pi} \int_{0}^{2\pi} |T_1(\phi,\theta,k r)|^2 sin\theta d\theta d\phi } \sqrt{\int_{0}^{\pi} \int_{0}^{2\pi}|T_2(\phi,\theta,k r)|^2 sin\theta d\theta d\phi}|^2}.
\label{eq:MSCdir}
    \end{aligned}
\end{equation}

Moreover, the general expression of the directivity of a first-order differential microphone is given by the following relationship:

\begin{equation}
	\begin{aligned}
	T_i(\psi_i) = \alpha_i + (1 - \alpha_i) \cos{\psi_i},
	\end{aligned}
\end{equation}

where $i \in [1,2]$ is the microphone index, $\psi_i$ is the angle between wave incidence and microphone orientation axis, and $\alpha_i \in [0,1]$ is the directivity parameter of the microphone $i$, which ranges from bidirectional ($\alpha_i = 0$) to omnidirectional ($\alpha_i = 1$). \\

For first-order differential microphones, there is a closed-form expression for the numerator and denominator of Eq.~(\ref{eq:MSCdir}):
\begin{equation}
	\begin{aligned}
    &N_{12}(k r) =  \frac{\alpha_1 \alpha_2 sin(kr)}{kr} 
    + \frac{(1-\alpha_2)(1-\alpha_2)(x_1x_2+y_1y_2)}{(kr)^3}(sin(kr)-kr cos(kr)) \\
    &+ \frac{z_1 z_2}{kr^3}[ ( (kr)^2 sin(kr) + 2kr cos(kr) )(1-\alpha_1)(1-\alpha_2) 
    + 2 sin(kr)(1-\alpha_1)(1-\alpha_2) ] \\
    &+ \frac{z_1}{(kr)^3}[ j(kr)^2 \alpha_2 cos(kr)(\alpha_1-1) + jkr \alpha_2 sin(kr)(1+\alpha_1) ] \\
    &+ \frac{z_2}{(kr)^3}[ j(kr)^2 \alpha_1 cos(kr)(\alpha_2-1) + jkr \alpha_1 sin(kr)(1+\alpha_2) ],\\
    &D_{12}(kr) =  \frac{\sqrt{3 \alpha_1^2+(1-\alpha_1)^2}\sqrt{3 \alpha_2^2+(1-\alpha_2)^2}}{3},
    \label{eq:closedform_msc}
    \end{aligned}
\end{equation}
where: $ x_i = cos(\phi_i) sin(\theta_i); y_i = sin(\phi_i) sin(\theta_i); z_i = cos(\theta_i)$
refers to the wave incidence angle $\psi_i$ expressed in spherical coordinates (with azimuth $\phi$ and inclination $\theta$).

\subsection{\label{subsec:3:2}Diffuseness estimation in ambisonics}
Let us consider a sound field captured with a spherical microphone array,
which contains $Q$ microphones distributed around a spherical surface of radius $R$
at the angular positions given by the azimuth-inclination pairs
$\Omega_q = (\phi_q,\theta_q)$.
The captured frequency-domain signals $ X_q(k)$ can be represented as the spherical harmonic domain signals $X_{mn}(k)$ through the spherical harmonic transform of order $L$ \cite{bertet_3d_2006}:
\begin{equation}
        X_{mn}(k) = \sum^{Q} X_q(k) Y_{mn}(\Omega_q) \Gamma_m(kR),
    \label{eq:ambisonics_encoding}
\end{equation}
where $Y_{mn}(\Omega_q)$ are the \textit{real-valued spherical harmonics}, and $\Gamma_m(kR)$ are the
\textit{radial filters} or equalization terms of order $m$, with $m \in [0,L]$ and $n \in [-m,m]$.\\

Due to a number of practical reasons, it is desirable to distribute the microphone capsules in a uniform manner along the sphere, with the the regular tetrahedron being the simplest possible configuration
 \cite{gerzon_design_1975}. Capsule signals recorded with such topology receive the name of \textit{A-Format} signals. Conversely, the term \textit{B-Format} (\textit{ambisonics}) describes the application of Eq.~(\ref{eq:ambisonics_encoding}) (\textit{ambisonic encoding}) to the \textit{A-Format} signals.
One of the most common coherence estimators for first-order ambisonic frequency-domain signals $X_{mn}(k)$ is the \textit{diffuseness} $\Psi$ as defined in \textit{DirAC} \cite{pulkki_directional_2006}:

\begin{equation}
    \Psi(k) = 1 - \frac{2 ||\left\langle \mathbb{R}\{\bm{X_1}(k)X_0(k)^* \} \right\rangle|| }{ \left\langle||\bm{X_1}(k)||^2 + |X_0(k)|^2\right\rangle},
    \label{eq:psi}
\end{equation}
where $X_0(k) = X_{00}(k)$ 
and $\bm{X}_1(k) = [X_{1-1}(k), X_{10}(k), X_{11}(k)]^\intercal$ are \textit{SN3D}-normalized.
For the sake of clarity, we will further define the \textit{B-Format coherence} estimator $\Delta$ as:
\begin{equation}
	\Delta(k) = 1 - \Psi(k).
	\label{eq:delta}
\end{equation}

Under spherical isotropic noise, the theoretical coherence between any pair of zeroth and first order ambisonic virtual microphones is equal to 0 for all frequencies, due to orthogonality and symmetry of the spherical harmonics \cite{elko_spatial_2001}. . This result can be also assessed by Eq.~(\ref{eq:closedform_msc}).

However, there are several practical factors that might corrupt the coherence estimation, such as the approximation of the temporal expectation by time averaging \cite{thiergart_diffuseness_2011} in Eq.~(\ref{eq:psi}), or the non-ideal implementation of the radial filters $\Gamma_m(kR)$ \cite{schorkhuber_ambisonic_2017}.
In the following sections, we present several experiments that illustrate the behavior of different coherence estimators applied on the signals captured with a tetrahedral microphone subjected to spherical isotropic noise, using both simulated and real sound recordings.

\section{Methods}

\subsection{Simulation}
Spherical isotropic noise has been generated following the \textit{geometrical method} \cite{habets_generating_2007, habets_comments_2010}, using $N = 1024$ plane waves. The resulting \textit{A-Format} signals correspond to a virtual tetrahedral microphone array mimicking the Ambeo
\footnote{Sennheiser Ambeo VR Mic. 
\todo{https://en-us.sennheiser.com/microphone-3d-audio-ambeo-vr-mic}}
characteristics ($R=0.015$ meter, $\alpha=0.5$). 
The generated audio has a duration of 60 seconds.

\subsection{Recording}
Spherical isotropic noise has been rendered to a spherical loudspeaker layout with 25 \textit{Genelec 8040}. The loudspeakers are arranged into three azimuth-equidistant 8-speaker rings at inclinations $\theta = [\pi/4, \pi/2, 3\pi/4]$, plus one speaker at the zenith ($\theta=0$).
The different speaker distances to the center are delay- and gain-corrected, and the signal feeds are equalized to compensate for speaker coloration. The room has an approximate $T_{60}$ of 300 ms measured at the 1 kHz third-band octave. 
The spherical isotropic noise has been again created following the \textit{geometrical method}, encoding a number of uncorrelated noise plane waves in ambisonics with varying orders $L \in [1,5]$. Due to practical limitations related with the software, the minimum number of sources $N = 256$ for an accurate sound field reconstruction \cite{habets_comments_2010} could not be reached - instead, the analysis has been performed parametrically with $N = [8, 16, 32, 64]$.
For each value of $L$ and $N$, approximately 15 seconds of audio have been recorded with an Ambeo microphone located at the center of the speaker array.
Ambisonics decoding uses the AllRAD method \cite{zotter_all-round_2012}, passing through a spherical 64-point 10-design virtual speaker layout, and includes an imaginary speaker at the nadir ($\theta=\pi$). The decoding matrix uses \textit{in-phase} weights.

\subsection{Data processing and metrics}

The sampling rate of all signals is 48 kHz.
All frequency-domain results have been obtained by averaging their time-frequency representations over time.  
Ambisonics conversion is performed using \textit{Ambeo A-B converter} AU plugin, version 1.2.1.

Two error metrics are considered: the frequency-dependent squared error $\varepsilon(k)$, and the mean squared error $\bar{\varepsilon}$:

\begin{equation}
    \varepsilon(k) = |X_1(k) - X_2(k)|^2; \text{         }\bar{\varepsilon} = \frac{1}{K}{\sum_{k=1}^{K} |X_1(k) - X_2(k)|^2}
    \label{nmse}
\end{equation}

\section{Results and discussion}
\subsection{\label{subsec:results_aformat}A-Format}

\begin{figure}
	\includegraphics[width=\textwidth]{Figures/CoherenceEstimation/Figure1}
    \caption{\label{fig:Fig1}\textit{A-Format} coherence between microphone signals. Left: $\text{MSC}$ as a function of the frequency of theoretical, simulated and recorded \textit{((BLD,BRU)},  $L=5, N=64$) signals. Right: mean error $\bar{\varepsilon}$ of the recorded signals' $MSC$ \textit{(BLD,BRU)} compared to the simulated values, for all values of $L$ and $N$.}
\end{figure}

The coherence of the generated \textit{A-Format} signals is exemplified in Fig.~\ref{fig:Fig1} (left), which shows the $MSC$ between the capsule pair (\textit{BLD,BRU}) for the theoretical, simulated and recorded cases.
The theoretical coherence is derived from Eq.~(\ref{eq:closedform_msc}), while simulated and recorded MSC have been computed by Welch's method, using a \textit{hanning} window of 256 samples and 1/2 overlap.
The difference between theoretical and simulated coherence is negligible for practical applications.
However, there is a noticeable difference when compared to the recorded coherence. 
In general, the recorded $\text{MSC}$ follows the tendency of the simulated curve up to around 5 kHz.
Above this frequency, the recorded $MSC$ presents several spectral peaks, which might be partially explained by the interference of the microphone itself in the recorded sound field, and by the non-ideal directivity of the capsules.
The squared error $\varepsilon(k)$ with respect to the simulated curve is shown in Fig.~\ref{fig:Fig1} (left), while Fig.~\ref{fig:Fig1} (right) represents the same error averaged over frequency $\bar{\varepsilon}$ for different spatial resolution values of the diffuse field reproduction algorithm.
As expected, $\bar{\varepsilon}$ decreases with increasing values of $L$ and $N$.

\subsection{B-Format} 
\begin{figure}
	\includegraphics[width=\textwidth]{Figures/CoherenceEstimation/Figure2}
	\caption{\label{fig:Fig2} Estimated \textit{B-Format} coherence ($\Delta$) of a simulated diffuse sound field, as a function of the temporal averaging vicinity radius $r$. Left: $\Delta(k)$ for different values of $r$, with (coarse) and without (fine) application of radial filters. Right: mean and standard deviation of $\Delta(k)$ as a function of $r$.}
\end{figure}

In order to evaluate the dependency of $\Delta$ on the number of time frames used for averaging, the following procedure is presented.
The simulated \textit{A-Format} sound field has been transformed into the spherical harmonic domain, with and without the application of radial filters $\Gamma_m(kR)$. Then, $\Delta$ has been computed with Eq.~(\ref{eq:delta}) for exponentially growing values of $r$ between 1 (8 ms) and 2048 (10.92 s), where $r$ is the vicinity radius used for time averaging, and the number of time windows is given by $T = 2r+1$.
The time-frequency representation is derived by applying the STFT with the same window parameters as in Subsection \ref{subsec:results_aformat}.\\

Figure~\ref{fig:Fig2} (left) shows the great dependence of $\Delta$ on $r$.  The estimated coherence tends to the theoretical values with increasing values of $r$. This tendency is better appreciated in Fig.~\ref{fig:Fig2} (right): the curve asymptotically decreases to a value $\Delta_{min}\approx0$.
Another interesting observation comes from the frequency response of the curves. For all values of $r$, the coherence of the compensated \textit{B-Format} signal (with $\Gamma_m(kR)$) is roughly flat up to around 7 kHz, which approximately corresponds to the operational spatial frequency range of the microphone \cite{gerzon_design_1975}.
Above this value, the coherence response looses the flatness due to spatial aliasing. The response above the maximum frequency could be stabilized, if needed, by alternative diffuseness estimation methods \cite{politis_direction--arrival_2015}.
The coherence level differences along frequency are inversely proportional to $r$ --- the effect is better depicted by the standard deviation values (right).
The effect of the radial filters in the coherence measurement is also shown: for a given $r$, the shape of the coherence is always less flat if no filters are applied. Conversely, in this case, coherence values are always smaller for the same $r$. This effect might be explained taking into account the inter-channel coherence introduced by microphone and encoder imperfections in real scenarios \cite{schorkhuber_ambisonic_2017}.
As a remark, the comparison between Figs.~\ref{fig:Fig1} and \ref{fig:Fig2} provides evidence that the application of the spherical harmonic transform might be able to yield more accurate diffuseness estimations, due to a better signal conditioning \cite{epain_spherical_2016}.\\

\begin{figure}
	\includegraphics[width=\textwidth]{Figures/CoherenceEstimation/Figure3}
	\caption{\label{fig:Fig3}\textit{B-Format} coherence between microphone signals. Left: $\Delta$ of simulated and recorded ($L=5, N=64$) signals. Right: $\bar{\varepsilon}$ of the recorded signals coherence across all values of $L$ and $N$.}
\end{figure}

Figure~\ref{fig:Fig3} (left) shows the estimated coherence for the recorded sound field with $L=5$ and $N=64$, using a vicinity radius of $r=1024$ ($\approx$ 5 s).
The curve is centred around $\Delta=0.25$ and presents several spectral peaks, as in the \textit{A-Format} case. 
It is important to notice here that the deviations between the coherence of the simulated and the recorded sound fields are much stronger compared to those of Fig.~\ref{fig:Fig1}. 
This effect can be also appreciated in Fig.~\ref{fig:Fig3} (right): the mean squared error is around two orders of magnitude higher in \textit{B-Format}.
Nevertheless, similar as in Fig.~\ref{fig:Fig1} (right), $\bar{\varepsilon}$ decreases with increasing values of $L$ and $N$.
This behavior suggests that the deviations between the recorded and the simulated coherence can be to a large degree explained by the low spatial resolution of the reproduction system; given a higher number of loudspeakers, we expect that the reproduced diffuseness will tend to the theoretical expression.

\section{Conclusions}
The diffuseness of a sound field is an important parameter for several applications. In this work, two different metrics of diffuseness have been defined and measured with a tetrahedral microphone subjected to spherical isotropic noise.
The analysis shows, first, the impact of the time-averaging window length on the \textit{B-Format} diffuseness estimator.
This result might be useful for designing coherence estimators that are parametrized with respect to the length of the analysis window \cite{thiergart_diffuseness_2011}.
Second, the feasibility of diffuse sound field reproduction by a spherical loudspeaker array using ambisonics plane-wave encoding and the \textit{geometrical method} is studied. 
Results suggest that this approach is viable, given a sufficient spatial resolution; a quantification of the impact of the number of loudspeakers remains for future work.