### 2020-06-01 Chapters/Conclusions.tex
update 20200602

\chapter{Conclusions}

\section{Summary of Contributions}

\begin{itemize}
  \item Academic Contributions

  \begin{enumerate}
  	\item Analysis of spherical isotropic noise fields with an A-Format tetrahedral microphone \todo{ref}
  	\begin{itemize}
  		\item \textbf{Parameter estimation}: Contribution to the characterization of coherence with tetrahedral microphones (the most common spherical arrangement)
	\end{itemize}

  	\item Blind reverberation time estimation from ambisonic recordings \todo{ref}
  	\begin{itemize}
  		\item \textbf{Parameter estimation}: Novel technique for blind RT60 estimation of ambisonic recordings from autoregressive models
	\end{itemize}

  	\item A hybrid parametric-deep learning approach for sound event localization and detection \todo{ref}
  	\begin{itemize}
  		\item \textbf{Scene Description}: Novel state-of-the-Art methodology for Sound Event Localization and Detection	
  	\end{itemize}

  \end{enumerate}

  \item Software Contributions

	\begin{enumerate}

  		\item Ambiscaper: A Tool for Automatic Generation and Annotation of Reverberant Ambisonics Sound Scenes \todo{ref}
  		\begin{itemize}
  			\item \textbf{Data Generation}: Novel tool for reverberant ambisonic dataset generation
		\end{itemize}

  		\item  Ambisonics Directional Room Impulse Response as a New Convention of the Spatially Oriented Format for Acoustics \todo{ref}
		\begin{itemize}
  			\item \textbf{Recorded IRs}: File standard/convention proposal for storage of recorded ambisonic IRs
		\end{itemize}

		\item A Python library for Multichannel Acoustic Signal Processing \todo{ref}
		\begin{itemize}
  			\item \textbf{Simulated IRs}: Library for acoustic simulation (IR generation, microphone array simulation, etc)
		\end{itemize}

  		\item pysofaconventions, a Python API for SOFA \todo{ref}
		\begin{itemize}
  			\item \textbf{Recorded IRs}: implementation of SOFA for python 
		\end{itemize}

	\end{enumerate}

\end{itemize}

\section{List of related publications}