### 2020-06-03 manuscript.tex

\documentclass[12pt, b5paper,twoside]{tesi_upf_new}
%\documentclass[12pt, b5paper,twoside]{book}

%CODIFICATION
\usepackage[latin1]{inputenc}

%LENGUAGE
\usepackage[english,catalan]{babel}

%ONLY TO OBTAIN MARK BANK INDEX INDICATION B5
\usepackage[b5paper]{geometry}
\usepackage[cam,a4,center,frame]{crop}

%INCLUDE GRAPHICS AND THE LOGO OF THE UPF
\usepackage{graphicx}

%FONTS TIMES OR GARAMOND, 
\usepackage{times}
%\usepackage{garamond}

%COHERENCE ESTIMATION
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{latexsym}% 
\usepackage{lineno}   
\usepackage{array}
\usepackage{adjustbox}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{newpxtext} 
\usepackage{cite}
\usepackage{booktabs}

% COMMENT THESE FOR TESI_UPF
\usepackage{hyperref}
\hypersetup{pdftex,colorlinks=true,allcolors=blue}
\usepackage{hypcap}

%SELD
\DeclareMathOperator{\LL}{\langle}
\DeclareMathOperator{\RR}{\rangle}
\usepackage{graphicx,url,times,booktabs, tabularx}
\usepackage{array}
\usepackage{booktabs}
\usepackage{algorithm, algorithmic}
\usepackage{color}
\usepackage{subcaption}

%\usepackage{hyperref}

%WITHOUT HEADINGS: NO MODIFICATION
\pagestyle{plain}

%FOR THE INDEX OF SUBJECTS
\usepackage{makeidx}
\makeindex

%BIBLIOGRAPHY STYLE
\bibliographystyle{apalike}

%SELECT LANGUAGE
\selectlanguage{english}

%THE TABLE OF CONTENTS IS TITLE CONTENTS
%\addto\captionscatalan
  {\renewcommand{\contentsname}{\Large \sffamily Contents}}

%ADD YOUR DATA
\title{Ambisonic domain methods for immersive audio production enhancement}
%\subtitle{\todo{subtitle Required?}}
\author{Author: Andrés Pérez López}

% TODO UNCOMMENT FOLOWING
\thyear{2020}
\department{of Information and Communication Technologies}
\supervisor{Dr. Adan Garriga Torres\\Dr. Emilia Gómez Gutiérrez}

% TODO
\newcommand\todo[1]{\textcolor{red}{#1}}

% INTEGRAL D
\newcommand{\dif}{\mathop{}\!\mathrm{d}}

\begin{document}

\frontmatter

\maketitle

\cleardoublepage

%%%%%% Dedication

\noindent To my cat

\cleardoublepage

%%%%%% End dedication

%%%%%% Thanks
\noindent {\Large \sffamily Thanks} thanks to....  \todo{thanks}

\cleardoublepage

%%%%%% End of thanks

%ABSTRACT IN TWO LEGUAGES. 150 words max!!!
\selectlanguage{english}
\section*{\Large \sffamily Abstract}

Recent years have witnessed a rise in immersive multimedia content, as a result of the developments in Virtual and Augmented Reality, where spatial audio plays a crucial role for the creation of a realistic experience. 
Ambisonic microphones allow the recording of sound scenes while preserving their spatial characteristics. For that reason, ambisonic has turned into the \textit{de facto} standard for immersive audio transmission. Yet, few production tools exploit all the information available in such media.   
In the present thesis we raise the research question of obtaining relevant information from ambisonic recordings, in a way that is meaningful for producers.
In order to answer the question, we have developed several methods and tools applied to different levels of the immersive sound production workflow; with each of those elements representing a novel contribution to the state-of-the-art in their respective fields.

\selectlanguage{catalan}
\vspace*{\fill}
\section*{\Large \sffamily  Resum}

En els últims anys hem assistit a l'auge del contingut multimèdia immersiu, afavorit pels avanços en Realitat Virtual i Augmentada. En aquest mitjà, el so espacial juga un paper molt important per a la creació d'una sensació de realisme.
Els micròfons ambisonics permeten gravar escenes sonores preservant les seues característiques espacials. A causa d'això, ambisonics s'ha convertit en l'estàndard de facto per a la transmissió d'àudio immersiu. No obstant això, encara son poques les eines de producció que permeten explotar tota la informació disponible en aquest mitjà.
En la present tesi plantegem la pregunta de com obtindre informació rellevant de gravacions ambisonics de manera automàtica, d'una forma que siga significativa per als productors.
Per a això, hem desenvolupat una sèrie de mètodes i eines que s'apliquen a diferents nivells de la cadena de producció de so immersiu; cadascun d'aquests elements representa una contribució nova a l'estat de l'art en aquest camp.

%En los últimos años hemos asistido al auge del contenido multimedia immersivo, auspiciado por los avances en Realidad Virtual y Aumentada. En este medio, el sonido espacial juega un papel muy importante para la creación de una sensación de realismo.
%
%Los micrófonos ambisonics permiten grabar escenas sonoras preservando sus características espaciales.
%Debido a esto, ambisonics se ha convertido en el estándar de facto para la transmisión de audio immersivo. Sin embargo, todavía hay pocas herramientas de producción que permitan explotar toda la información disponible en este medio.
%
%En esta tesis planteamos la pregunta de cómo obtener información relevante de grabaciones ambisonics de manera automática, de una forma significativa para los productores.
%
%Para ello, hemos desarrollado una serie de métodos y herramientas que se aplican a diferentes niveles de la cadena de producción de sonido immersivo; cada uno de estos elementos representa una contribución novedosa al estado del arte en ese campo.

\vspace*{\fill}

\cleardoublepage
\selectlanguage{english}
%END OF ABSTRACT

%PREFACE. 
{\bf Preface \todo{is that really needed?}}

\cleardoublepage
%END OF PREFACE

%TABLE OF CONTENTS: REQUIRED
\tableofcontents

%lIST OF FIGURES; ONLY IF THERE ARE FIGURES
\listoffigures
%TO APPER THE LIST OF FIGURES IN THE TABLE OF CONTENTS 
\addcontentsline{toc}{chapter}{List of figures}

%LIST OF TABLES; ONLY IF THERE ARE TABLES
\listoftables
%TO APPEAR THE LIST OF TABLES IN THE TABLE OF CONTENTS
\addcontentsline{toc}{chapter}{List of tables}

%START THE TEXT
\mainmatter

%INSERT CHAPTERS
\input{Chapters/Introduction.tex}
\input{Chapters/ScientificBackground.tex}
\input{Chapters/CoherenceEstimation}
\input{Chapters/ReververationTimeEstimation}
\input{Chapters/SELD}
\input{Chapters/DataGeneration}
\input{Chapters/Conclusions}

%ADD BIBLIOGRAPHY
\bibliography{Bibliography/CoherenceEstimation, Bibliography/SELD.bib, Bibliography/Ambiscaper.bib, Bibliography/AmbisonicsDRIR.bib, Bibliography/ReverberationTimeEstimation.bib, Bibliography/Introduction.bib}

\backmatter
\printindex

\end{document}

%NUMBER OF THE EXTERNAL PAGE EXCEPT IN THE FIRST PAGE OF EACH CAPITAL
%\usepackage{fancyhdr}
%\pagestyle{fancy}
%\fancyfoot{}
%\fancyfoot[RO]{\thepage}
%\fancyfoot[LE]{\thepage}
%
%
%%MULTIPLE INDEX
%%In the preamble
%\usepackage{multind}
%\makeindex{authors}
%%Introduction to form entries
%\index{authors}{Einstein}
%%Situation of the Index
%\printindex{authors}{Author index}
%%The \ usepakage {makeidx} \ makeindex \ printindex commands must be removed
%%You need to exacute from the command line makeindex authors