### 2020-07-06 manuscript.tex
minor

\documentclass[12pt, b5paper,twoside]{tesi_upf_new}
%\documentclass[12pt, b5paper,twoside]{book}

%CODIFICATION
\usepackage[latin1]{inputenc}

%LENGUAGE
\usepackage[english,catalan]{babel}

%ONLY TO OBTAIN MARK BANK INDEX INDICATION B5
\usepackage[b5paper]{geometry}
\usepackage[cam,a4,center,frame]{crop}

%INCLUDE GRAPHICS AND THE LOGO OF THE UPF
\usepackage{graphicx}

%FONTS TIMES OR GARAMOND, 
\usepackage{times}
%\usepackage{garamond}

%COHERENCE ESTIMATION
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{latexsym}% 
\usepackage{lineno}   
\usepackage{array}
\usepackage{adjustbox}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{newpxtext} 
\usepackage{cite}
\usepackage{booktabs}

% COMMENT THESE FOR TESI_UPF
\usepackage{hyperref}
\hypersetup{pdftex,colorlinks=true,allcolors=blue}
\usepackage{hypcap}

%SELD
\DeclareMathOperator{\LL}{\langle}
\DeclareMathOperator{\RR}{\rangle}
\usepackage{graphicx,url,times,booktabs, tabularx}
\usepackage{array}
\usepackage{booktabs}
\usepackage{algorithm, algorithmic}
\usepackage{color}
\usepackage{subcaption}

\usepackage[none]{hyphenat}
\tolerance=1
\emergencystretch=\maxdimen
\hyphenpenalty=10000
\hbadness=10000

\usepackage[titletoc]{appendix}

%WITHOUT HEADINGS: NO MODIFICATION
\pagestyle{plain}

%FOR THE INDEX OF SUBJECTS
\usepackage{makeidx}
\makeindex

%BIBLIOGRAPHY STYLE
\bibliographystyle{apalike}

%SELECT LANGUAGE
\selectlanguage{english}

%THE TABLE OF CONTENTS IS TITLE CONTENTS
%\addto\captionscatalan
  {\renewcommand{\contentsname}{\Large \sffamily Contents}}

%ADD YOUR DATA
%\title{Ambisonic domain methods for immersive audio production enhancement}
\title{Parametric analysis of ambisonic audio}
\subtitle{Contributions to methods, applications and data generation}
\author{Author: Andrés Pérez López}

% TODO UNCOMMENT FOLOWING
\thyear{2020}
\department{of Information and Communication Technologies}
\supervisor{Dr. Adan Garriga Torres\\Dr. Emilia Gómez Gutiérrez}

% TODO
\newcommand\todo[1]{\textcolor{red}{#1}}

% INTEGRAL D
\newcommand{\dif}{\mathop{}\!\mathrm{d}}

\begin{document}

\frontmatter

\maketitle

\cleardoublepage

%%%%%% Dedication

\noindent To my cat

\cleardoublepage

%%%%%% End dedication

%%%%%% Thanks
\noindent {\Large \sffamily Thanks}\\

First of all, I must thank my supervisors, Emilia and Adan, who have made possible the realization of this journey.

Thanks to all my friends from the Audio Team at Eurecat: Umut, Julien, Tim, Niklas, Toni, Gerard, Andrés, John, Stefan, Marc. And also to  the people from Image and the Visualization teams. I spent last four years working with the finest of the companies.

Thanks also to all the great people I met at the MTG, including Pritish, Xavier, and specially Edu for the logistic and intellectual help. I would like to give public acknowledgment to Xavier Serra for his successful task of leading and gathering such unique international group of talented people.

Special thanks to Nikos and all the Signal Processing Lab crew for warmly welcoming me in Heraklion, and helping me in many different ways in your lovely place. I hope to come back there soon.

Many thanks to Archontis and Tuomas for the visit opportunity, and for all the  facilities they gave me in my stay in this amazing place called Tampere. I take the opportunity to say hello and thanks to Kostas, Natalya, Annamaria, Toni, and all the nice people there.

Thanks to Max, who has been always very receptive and supportive on numerous occasions, and has indirectly contributed to the success of this project. 
Thanks also to Rafa, who decided to get on board for some unknown reason.

Thanks to all the friends I made and preserved in Barcelona in the last four years. Special thanks to my musician family from Balkumbia, Fanfarrai and Telewawachi Kilili; I learned a lot from you. Special acknowledgements for the good times with the people in Can Collet (best farm-school in town) and the Valencian oldies.

Thanks a lot to Chedar for his company in the final steps. And, of course, million thanks to my family and to Marta for their unconditional love and support, powering all the way up here. \\

To all of you: thanks.

\cleardoublepage

%%%%%% End of thanks

%ABSTRACT IN TWO LEGUAGES. 150 words max!!!
\selectlanguage{english}
\section*{\Large \sffamily Abstract}

Due to the recent advances in virtual and augmented reality, ambisonics has emerged as the \textit{de facto} standard for immersive audio.
Ambisonic audio can be captured using spherical microphone arrays, which are becoming increasingly popular.
Yet, many methods for acoustic and microphone array signal processing are not specifically tailored for spherical geometries.
Therefore, there is still room for improvement in the field of automatic analysis and description of ambisonic recordings. 
In the present thesis, we tackle this problem using methods based on the parametric analysis of the sound field. 
Specifically, we present novel contributions in the scope of blind reverberation time estimation, diffuseness estimation, and sound event localization and detection. 
Furthermore, several software tools developed for ambisonic dataset generation and management are also presented.

\selectlanguage{catalan}
\vspace*{\fill}
\section*{\Large \sffamily  Resum}

\textit{Ambisonics} ha esdevingut l'estàndard d'àudio immersiu als últims anys, afavorit pels avançaments en realitat virtual i augmentada.
L'àudio \textit{ambisonic} es pot obtindre mitjançant \textit{arrays} de micròfons esfèrics, que són cada vegada més populars.
Tot i això, la majoria de mètodes acústics i de processament de senyal basats en \textit{arrays} de micròfons no estan adaptats al cas específic de geometries esfèriques.
Per tant, encara hi ha moltes possibilitats de millora en l'àmbit d'anàlisi automàtica i descripció de gravacions \textit{ambisonic}.
En la present tesi plantegem aquest problema basant-nos en l'anàlisi paramètrica del camp acústic.
Més concretament, presentem contribucions originals en les àrees d'estimació de reverberació, estimació de difusió acústica, i detecció i localització d'esdeveniments sonors.
Així mateix, presentem diverses eines de programari desenvolupades per la generació i manteniment de bases de dades d'àudio \textit{ambisonic}.

\vspace*{\fill}

\cleardoublepage
\selectlanguage{english}
%END OF ABSTRACT

%PREFACE. 
\section*{\Large \sffamily  Preface}

The present thesis has been carried out within the context of the Industrial Doctorate program from the Catalan government, as a collaboration between the Music Technology Group of the Pompeu Fabra University, and the Multimedia Unit of Eurecat, the Catalan Technology Center. 
It has been jointly supervised by Dr. Emilia Gómez and Dr. Adan Garriga.

The industrial nature of the doctorate program implies a tendency towards the innovative application of the findings and conclusions generated during the development of the thesis. 
Therefore, one of the targets of the research performed in the context of this thesis is the stress on real-world problems, scenarios and technologies, with the aim of developing methods that could be easily adapted into applications with a high level of technology readiness. 
Moreover, there has been also an stress on research reproducibility. This has not been motivated only by an academic perspective; but also by the open research and open innovation paradigms present in our today's society.\\

The work presented here would not have been possible without the contributions from an excellent network of researchers. Apart from internal collaborations (UPF and Eurecat), two of the main contributions in the thesis followed two research visits to internationally recognized institutions: the Audio Research Group from Tampere University, Finland (Chapter~\ref{chap:rt60}), and the Signal Processing Group at the FORTH - Foundation for Research and Technology Hellas, Heraklion, Greece, (Chapter~\ref{chap:coherence}).

The alphabetical list of academic collaborators is: Eduardo Fonseca, Emilia G\'omez, Rafael Ib\'añez-Usach,  Julien de Muynke, Archontis Politis, Xavier Serra, and Nikolaos Stefanakis.

\cleardoublepage
%END OF PREFACE

%TABLE OF CONTENTS: REQUIRED
\tableofcontents

%lIST OF FIGURES; ONLY IF THERE ARE FIGURES
\listoffigures
%TO APPER THE LIST OF FIGURES IN THE TABLE OF CONTENTS 
\addcontentsline{toc}{chapter}{List of figures}

%LIST OF TABLES; ONLY IF THERE ARE TABLES
\listoftables
%TO APPEAR THE LIST OF TABLES IN THE TABLE OF CONTENTS
\addcontentsline{toc}{chapter}{List of tables}

%START THE TEXT
\mainmatter

%INSERT CHAPTERS
\input{Chapters/Introduction.tex}
\input{Chapters/ScientificBackground.tex}
\input{Chapters/ReververationTimeEstimation}
\input{Chapters/CoherenceEstimation}
\input{Chapters/SELD}
\input{Chapters/DataGeneration}
\input{Chapters/Conclusions}

%ADD BIBLIOGRAPHY
\bibliography{Bibliography/CoherenceEstimation, Bibliography/SELD.bib, 
	Bibliography/Data.bib, 
	Bibliography/AmbisonicsDRIR.bib, 
	Bibliography/ReverberationTimeEstimation.bib,
	Bibliography/Introduction.bib, 
	Bibliography/ScientificBackground.bib,
	Bibliography/Conclusion.bib}

\backmatter
\printindex

\end{document}

%NUMBER OF THE EXTERNAL PAGE EXCEPT IN THE FIRST PAGE OF EACH CAPITAL
%\usepackage{fancyhdr}
%\pagestyle{fancy}
%\fancyfoot{}
%\fancyfoot[RO]{\thepage}
%\fancyfoot[LE]{\thepage}
%
%
%%MULTIPLE INDEX
%%In the preamble
%\usepackage{multind}
%\makeindex{authors}
%%Introduction to form entries
%\index{authors}{Einstein}
%%Situation of the Index
%\printindex{authors}{Author index}
%%The \ usepakage {makeidx} \ makeindex \ printindex commands must be removed
%%You need to exacute from the command line makeindex authors